Current and future constraints on primordial magnetic fields

### 2017-01-23 pmf.tex
rnd 1 thru forecasts

@@

\begin{figure}[htb]\centering
\includegraphics[width=0.45\textwidth,clip,trim={1.5cm 12.5cm 11cm 7.5cm}]{pmf_area.pdf}
  \caption[Area dependence]{
  The forecasted uncertainty on the PMF power, \apmf{}, falls as the survey area increases. 
  In particular, observing more than $\sim$\,15\% of the sky is strongly preferred for PMF searches. 
  Fortunately, the CMB-S4 survey concepts being discussed cover even larger areas of sky. 
  The black solid line shows the forecast uncertainty as a function of area in the 11-parameter cosmological model.
  The purple dot-dash line shows the same, except in the restricted \lcdm{}+\apmf{} 7-parameter model. 
  In both cases, better limits on PMFs will be achieved with wider surveys.
    \label{fig:area}
  }
\end{figure}
A second question is whether it is better to integrate deeply on a small patch of sky or observe a wide area. 
As shown in  Fig.~\ref{fig:area}, observing at least 15\% of the sky is very important to a PMF search, and it is best  to cover as much sky as possible. 
A caveat to this analysis is that it is likely easier to remove galactic foregrounds to a specified level on targeted, `clean' patches as opposed to a substantial fraction of the sky, and we would expect galactic foregrounds to be important for the large angular scales. 
With that said, the PMF constraints improve by a factor of 1.4-1.5 (depending on the cosmological model) going from 10\% to 25\% of the full sky, and improve by another factor of 1.4-1.5 by going to the 70\% of the sky (the widest area likely to be possible after galactic cuts).

\tbd{got to here}

@@

Next we turn to the recovery of large angular scales, and the noise performance at low frequencies. 
We look at the impact of shifting the 1/f knee of the map-space noise. 
Effectively, we are multiplying the noise power, $N_\ell$, which is a constant for white noise,  by a function of angular multipole:
\be \label{eqn:knee}
f(\ell) = 1 + \left(\frac{\ell_{\rm knee}}{\ell}\right)^{8/3}.
\ee 
The exponent, 8/3, was selected based on a Kolmogorov spectrum of turbulence within a thin plane \citep{lay00}. %lay & halverson 2000. 
Note that this knob  serves as a placeholder for several effects, including a signal-to-noise hit due to galactic foregrounds or the methods used to clean these foregrounds, atmospheric noise, or actual instrumental 1/f noise. 
We find the results to be somewhat more sensitive to the 1/f knee in the simple \lcdm{}+\apmf{} model, where dropping \lknee{} from 400 to 50 reduces the uncertainty by a factor of 6.3.
However, the knee frequency is still important in the more conservative 11-parameter model, where the same shift reduces the uncertainties by a factor of 2.7. 
In both cases, as can be seen in Fig.~\ref{fig:knee}, the information loss accelerates once the 1/f knee goes above $\ell \simeq 200$. 
Recovering large angular scales is clearly important to PMF searches. 
This is consistent with the study of telescope size in the last paragraph - which found extending to smaller angular scales only helped slightly. 
It is also consistent with the current upper limits, which are dominated by the BICEP2/Keck Array data at low-$\ell$. 
As  these large angular scales are also crucial to searches for inflationary gravitational waves, hitting this target should not place new requirements on CMB-S4. 
The needs of the PMF search are well-aligned with the search for inflationary gravitational waves.

>>>>

\subsubsection{Telescope size}

@@

\subsubsection{Survey Area}

\begin{figure}[htb]\centering
\includegraphics[width=0.45\textwidth,clip,trim={1.5cm 12.5cm 11cm 7.5cm}]{pmf_area.pdf}
  \caption[Area dependence]{
  The forecasted uncertainty on the PMF power, \apmf{}, falls as the survey area increases. 
  In particular, observing more than $\sim$\,15\% of the sky is strongly preferred for PMF searches. 
  Fortunately, the CMB-S4 survey concepts being discussed cover even larger areas of sky. 
  The black solid line shows the forecast uncertainty as a function of area in the 11-parameter cosmological model.
  The purple dot-dash line shows the same, except in the restricted \lcdm{}+\apmf{} 7-parameter model. 
  In both cases, better limits on PMFs will be achieved with wider surveys.
  \tbd{may change strength of statements; haven't updated figure or caption}
    \label{fig:area}
  }
\end{figure}

\tbd{may change strength of statements}
A second question is whether it is better to integrate deeply on a small patch of sky or observe a wide area. 
Unsurprisingly, we find opposing preferences for the scale-invariant $n_b=-2.9$ (which peaks at large scales) and very blue $n_b=2$ spectrum peaking at small scales. 
We find very little dependence on area for the intermediate case of $n_b=-1$. 
As shown in  Fig.~\ref{fig:area}, for $n_b=-2.9$, observing at least 15\% of the sky is very important to a PMF search, and it is best  to cover as much sky as possible. 
A caveat to this analysis is that it is likely easier to remove galactic foregrounds to a specified level on targeted, `clean' patches as opposed to a substantial fraction of the sky, and we would expect galactic foregrounds to be important for the large angular scales. 
With that said, the PMF constraints improve by a factor of 1.4-1.5 (depending on the cosmological model) going from 10\% to 25\% of the full sky, and are nearly flat (a 6\% improvement) from 25\% to 70\% of the sky (the widest area likely to be possible after galactic cuts).
Conversely, $n_b=+2$, the optimal survey area is as small as possible. 
Note however that the minimum is extremely broad with only a 10\% worsening of the expected uncertainty as the survey area is increased from 10\% to 70\% of the  full sky. 
The optimal survey area for a PMF survey will depend on the targeted PMF models.

\subsubsection{Fidelity at large angular scales}

@@

Next we turn to the recovery of large angular scales, and the noise performance at low frequencies. 
We look at the impact of shifting the 1/f knee of the map-space noise. 
Effectively, we are multiplying the noise power, $N_\ell$, which is a constant for white noise,  by a function of angular multipole:
\be \label{eqn:knee}
f(\ell) = 1 + \left(\frac{\ell_{\rm knee}}{\ell}\right)^{8/3}.
\ee 
The exponent, 8/3, was selected based on a Kolmogorov spectrum of turbulence within a thin plane \citep{lay00}. %lay & halverson 2000. 
Note that this knob  serves as a placeholder for several effects, including a signal-to-noise hit due to galactic foregrounds or the methods used to clean these foregrounds, atmospheric noise, or actual instrumental 1/f noise.

We generally find the forecasts to be insensitive to the 1/f knee or galactic foreground removal at the default beam size of 4$^\prime$. 
It appears that even in the nearly scale-invariant case, the PMF constraint is being driven by smaller angular scales due to the the degeneracy with the tensor-to-scalar ratio at large angular scales. 
Without this degeneracy, i.e. in the simple \lcdm{}+\apmf{} model, dropping \lknee{} from 400 to 50 reduces the uncertainty by a factor of 3.2 for a nearly scale-invariant PMF. 
In this case, the degradation accelerates once the 1/f knee goes above $\ell \simeq 200$. 
This is consistent with the statement for the same simple model that the current upper limits are dominated by the BICEP2/Keck Array data at low-$\ell$. 
For the bluer PMF spectra ($n_b=-1$ or 2), the forecasts have nearly zero dependence on the knee frequency. 
These results are consistent with the study of telescope size in the last section - which found that recovering smaller angular scales with larger telescopes helped (if one didn't consider cost).
For a conservative 11-parameter model, our ability to separate a PMF from other physics is coming from sub-degree angular scales. 
An instrument built to precisely measure gravitational lensing of the CMB should do very well at searching for PMFs.

\subsubsection{Beam and calibration uncertainties}
